---
title       : Shining Light on Overfitting
subtitle    : Photovoltaic Pricing Trends and Projections
author      : torenunez
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## Regression

- Almost any set of data can be perfectly fitted with some high-order complex polynomial.   
  - But what is the true benefit and at what cost?  
  
  
We will be looking at the following data. 
  - Photovoltaic System Pricing Trends: Historical, Recent, and Near-Term Projections 2015 Edition
  - Which can be found [here](https://en.openei.org/datasets/dataset/photovoltaic-system-pricing-trends-historical-recent-and-near-term-projections-2015-edition) 

--- &twocol
## Let's take a look at some Photovoltaic Pricing Trends
*** =left
```{r initial plot, echo = FALSE, warning=FALSE}
InstallationYear <- seq(1998,2014,1)  
DollarsPerWatt <- c(12.4, 11.8, 10.7, 11.1, 11.1, 10.0, 9.4, 9.0, 9.1, 9.2, 8.8, 8.4, 7.1, 6.3, 5.4, 4.7, 4.3)
PriceData <- data.frame(InstallationYear, DollarsPerWatt)

library(ggplot2)
myTheme <- theme(
      axis.text.x = element_text(size = rel(1.25)),
      axis.text.y = element_text(size = rel(1.25)),
      title = element_text(size = rel(1.5), face = "bold")
)

    ggplot(PriceData, aes(x=InstallationYear, y=DollarsPerWatt)) +
      xlab("Installation Year") + ylab("Residential 2014 $/W") +
      geom_point(shape=19, size=2) +    # Use hollow circles
      #stat_smooth(method = "lm", formula = y ~ poly(x, input$order), size = 1) + 
      scale_y_continuous(breaks = seq(0,16,1)) + 
      ggtitle("Median Installed Price Trends over Time") + 
      myTheme

```
*** =right
Do you see a trend?

--- &twocol
## Let's just draw a straight line.
*** =left
```{r linear, echo = FALSE, warning=FALSE}

 ggplot(PriceData, aes(x=InstallationYear, y=DollarsPerWatt)) +
      xlab("Installation Year") + ylab("Residential 2014 $/W") +
      geom_point(shape=19, size=2) +    # Use hollow circles
      stat_smooth(method = "lm", formula = y ~ poly(x, 1), size = 1) + 
      scale_y_continuous(breaks = seq(0,16,1)) + 
      ggtitle("Median Installed Price Trends over Time") + 
      myTheme

```
*** =right
```{r linear R, echo = FALSE, warning=FALSE}
      MyLm <- lm(DollarsPerWatt ~ poly(InstallationYear, 1, raw=TRUE))
      MyR <- summary(MyLm)$r.squared
```
This gives us an R^2 of `r round(MyR,2)`.  
That's pretty good...   
Do you think we could do any better?  


--- &twocol
## How about a second order polynomial?
*** =left
```{r second, echo = FALSE, warning=FALSE}

 ggplot(PriceData, aes(x=InstallationYear, y=DollarsPerWatt)) +
      xlab("Installation Year") + ylab("Residential 2014 $/W") +
      geom_point(shape=19, size=2) +    # Use hollow circles
      stat_smooth(method = "lm", formula = y ~ poly(x, 2), size = 1) + 
      scale_y_continuous(breaks = seq(0,16,1)) + 
      ggtitle("Median Installed Price Trends over Time") + 
      myTheme

```
*** =right
```{r second R, echo = FALSE, warning=FALSE}
      MyLm2 <- lm(DollarsPerWatt ~ poly(InstallationYear, 2, raw=TRUE))
      MyR2 <- summary(MyLm2)$r.squared
```
This gives us an R^2 of `r round(MyR2,3)`.  
Is that better?  


--- &twocol
## How about a third order polynomial?
*** =left
```{r third, echo = FALSE, warning=FALSE}

 ggplot(PriceData, aes(x=InstallationYear, y=DollarsPerWatt)) +
      xlab("Installation Year") + ylab("Residential 2014 $/W") +
      geom_point(shape=19, size=2) +    # Use hollow circles
      stat_smooth(method = "lm", formula = y ~ poly(x, 3), size = 1) + 
      scale_y_continuous(breaks = seq(0,16,1)) + 
      ggtitle("Median Installed Price Trends over Time") + 
      myTheme

```
*** =right
```{r third R, echo = FALSE, warning=FALSE}
      MyLm3 <- lm(DollarsPerWatt ~ poly(InstallationYear, 3, raw=TRUE))
      MyR3 <- summary(MyLm3)$r.squared
```
This gives us an R^2 of `r round(MyR3,4)`.  
 

--- &twocol
## You get the idea
### 9th degree polynomial fit
*** =left
```{r ninth, echo = FALSE, warning=FALSE}

 ggplot(PriceData, aes(x=InstallationYear, y=DollarsPerWatt)) +
      xlab("Installation Year") + ylab("Residential 2014 $/W") +
      geom_point(shape=19, size=2) +    # Use hollow circles
      stat_smooth(method = "lm", formula = y ~ poly(x, 9), size = 1) + 
      scale_y_continuous(breaks = seq(0,16,1)) + 
      ggtitle("Median Installed Price Trends over Time") + 
      myTheme

```
*** =right
```{r ninth R, echo = FALSE, warning=FALSE}
      MyLm9 <- lm(DollarsPerWatt ~ poly(InstallationYear, 9, raw=TRUE))
      MyR9 <- summary(MyLm9)$r.squared
```
This gives us an R^2 of `r round(MyR9,9)`.  
  
- It's only marginally better than the linear fit. 
  - It is better... but is it worth it? At what cost?
- How is it trending at the end?  
  - It is going up! That's actually contrary to what the latest data would suggest...

--- 
## Conclusion

- This is a very simple example that illustrates how simple might be better.  
  - Check out the [Practical Machine Learning](https://www.coursera.org/learn/practical-machine-learning) to learn about the importance of: 
    - Loss of generality of the model.
    - Overfitting
    - Training and Testing Data Sets
    - And much more!


